---
title: "Reproducible documents"
author: "Mikkel Alnor"
format: html
code-fold: FALSE
theme: darkly   # cosmo bootswatch theme
highlight-style: arrow
---

# Setup code chuck

```{r setup, message=FALSE,warning=FALSE}
library(knitr)
library(tidyverse)
library(ggplot2)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))

small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)

tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
```

## About me

-   Owns many shrimps

-   From SDU, FGM, ATLAS

I started my PhD the 1th of *October*

I am taller than i am **wide**

## Simple code

```{r}
3 * 3
```

## Testing for Git

Then type out a sentence below the header with some random words, maybe about the weather.

## Showing the data

```{r, eval=FALSE}
post_meal_data
```

```{r,echo=T, results='hide'}
glimpse(post_meal_data)
head(post_meal_data)
```

## Tidy data

Each variable is a column; each column is a variable. Each observation is a row; each row is an observation. Each value is a cell; each cell is a single value.

mistakes in the data BMI is combined NA values id column

```{r,echo=T, results='hide'}
colnames(post_meal_data)
post_meal_data$id
```

## 9 - Basic data visualization

```{r}
#| label: fig-bmi-histo
#| fig-cap: "Distribution of BMI."

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram(bins = 10)
```

## Exercise: discrete plots

```{r}
#| label: fig-Discrete-plots
#| fig-cap: "Glykemi and Groups"

post_meal_data |>
  ggplot(aes(x = Group, fill = as.factor(glykemi))) +
  geom_bar() +
  labs(fill = "Glykemi")
```

## Side by side plots

```{r}
#| label: fig-bmi-glycemia
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap:
#|   - "Distribution of BMI."
#|   - "Number of those with glycemia."
#| layout-ncol: 2

ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram(bins = 10)

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## 9.11 Plotting two continuous variables

```{r}
#| label: fig-BMI-auc_pg
#| fig-cap: "BMI and auc_pg,"
#| fig-subcap:
#|  - "geom point"
#|  - "geom hex"
#|  - "geom point and geom smooth"
#| layout-ncol: 3

post_meal_data |>
  ggplot(aes(x = BMI, y = auc_pg)) +
  geom_point()

post_meal_data |>
  ggplot(aes(x = BMI, y = auc_pg)) +
  geom_hex()

post_meal_data |>
  ggplot(aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth(fill = "red4")
```

# **10 - Collaborating with GitHub**

\### Pushing changes to remote

\### pulling

# **11 - Data management and wrangling**

```{r}
post_meal_data |>
  select(Insulin10)

post_meal_data |>
  select(-Group)

post_meal_data |>
  select(Age, BMI, Weight)

post_meal_data |>
  select(starts_with("PG"))

post_meal_data |>
  select(ends_with("DXA"))

post_meal_data |>
  select(contains("0"))
```

## Renaming specific columns

```{r}
colnames(small_post_meal_data)

small_post_meal_data |>
  rename(
    Auc_glucose = auc_pg,
    Auc_insulin = auc_ins
  )

basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)
```

```{r}
post_meal_data |>
  select(contains("OGTT"))

post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )

post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename_with(~ gsub("..", ".minus", .x, fixed = TRUE))
```
## filtering data by row
```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")


small_post_meal_data |>
  filter(Group == "CTR" & BMI >= 25)
small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```
