---
title: "Reproducible documents"
author: "Mikkel Alnor"
format: html
code-fold: FALSE
theme: Morph   # cosmo bootswatch theme
highlight-style: tango
---

# Setup code chuck

```{r setup, message=FALSE,warning=FALSE}
library(knitr)
library(tidyverse)
library(ggplot2)
library(viridis)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
```

## About me

-   PhD student

-   Owns many shrimps

-   From SDU, FGM, ATLAS

I started my PhD the 1th of *October*

I am taller than i am **wide**

## Simple code

```{r}
3*3
```

## Testing for Git

Then type out a sentence below the header with some random words, maybe about the weather.

## Showing the data

```{r, eval=FALSE}
post_meal_data

```

```{r,echo=T, results='hide'}

glimpse(post_meal_data)
head(post_meal_data)

```

## tidy data

Each variable is a column; each column is a variable. Each observation is a row; each row is an observation. Each value is a cell; each cell is a single value.

mistakes in the data BMI is combined NA values id column

```{r,echo=T, results='hide'}
colnames(post_meal_data)
post_meal_data$id
```

## 9 - Basic data visualization

```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo

ggplot(post_meal_data,aes(x = BMI)) +
  geom_histogram(bins=10)+
    scale_fill_viridis(discrete=T)
```

## Exercise: discrete plots

```{r}
```{r}
#| fig-cap: "Group"
#| label: fig-group
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()

```

See @fig-___ above for a cool plot!


```


